Draft for project

### 2014-10-29 draft_project.tex
Ting er dritt, ting er bra. Vætta faen!

\subsection{Control of the spatial derivative}
This bound follows nicely the $L^1$-contractivity in that its derivation hinges on a clever choice of function to multiply (\ref{GPME}) with. Before proceeding, we define  $\Psi$ as the primitive of $\Phi$ with $\Psi(0) = 0$, i.e.
\begin{equation}
\label{phi_primitive}
	\Psi(s) = \int_0^s \Phi(u) ds,
\end{equation}

@@

\begin{obs}
We can go even further and deal with the last term on the right hand side: Assuming $f,\Phi(u) \in L^2(Q_T)$, then
\begin{equation*}
\iint_{Q_T} f\Phi(u)dxdt \leq \left(\iint_{Q_T} f^2 dxdt\right)^\frac{1}{2}\left(\iint_{Q_T} \Phi(u)^2 dxdt\right)^\frac{1}{2},
\end{equation*}
by Cauchy-Schwarz'. With the inequality $\sqrt{ab} \leq \frac{1}{4c}a + cb$ for $c>0$
\begin{equation*}
\iint_{Q_T} f\Phi(u)dxdt \leq \frac{1}{4c}\iint_{Q_T}f^2dxdt + c\iint_{Q_T}\Phi(u)^2dxdt.
\end{equation*}
Lastly, because of the homogeneous Dirichlet conditions, we use the Poincaré inequality and set $c$ so that
\begin{equation}
\label{using_Poincare}
\iint_{Q_T} f\Phi(u)dxdt \leq C\iint_{Q_T}f^2dxdt + \frac{1}{2}\iint_{Q_T}|\nabla \Phi(u)|^2dxdt,
\end{equation}
and the final inequality becomes
\begin{equation}
\frac{1}{2}\iint_{Q_T}|\nabla \Phi(u)|^2dxdt + \int_\Omega \Psi(u(T))dx \leq \int_\Omega \Psi(u_0)dx + C\iint_{Q_T}f^2dxdt.
\end{equation}
\end{obs}

\subsection{Control of the temporal derivative}
A similar, but perhaps a bit more involved argument can be used to make an estimate the temporal derivative.

@@

A particular choice for $\zeta$, that's of interest is $\zeta(t) = 1$. Then we end up with the estimate
\begin{equation}
\label{temporal_bound}
\iint_{Q_T} \Phi'(u) |\partial_t u|^2 dxdt \leq \iint_{Q_T}\left\{ -\partial_t f \Phi(u)   +Cf\partial_tf \right\}dxdt, 
\end{equation}
and in addition if $\Phi'(u)$ is bounded over all relevant values of $u$, e.g. given my the $L^\infty$-bound, then what we have here is an $L^2$-estimate of $\partial_t \Phi(u)$.

@@

	\item[Control of the temporal derivative:] Now we let $\zeta = \zeta(t)$ be a $C^1([0,T],[0,1])$ cutoff function so that $\zeta(0) = \zeta(T) = 1$. If we multiply (\ref{approximate_problem}) by $\zeta \partial_tw_n$ and integrate over $Q_T$ we get
	\begin{equation}
	\label{temporal_bound_start}
	\iint_{Q_T}\zeta \Phi'_n(u_n)|\partial_t u_n|^2 dxdt = \iint_{Q_T} \left\{ \zeta \partial_t w_n \Delta w_n + \zeta \partial_t w_n f_n \right\} dxdt.
	\end{equation}
	Let's start by considering the first term on the right hand side. By the product rule for the divergence
	\begin{equation*}
	\int_0^T \zeta \int_\Omega \partial_t w_n \Delta w_n dxdt = \int_0^T \zeta \int_\Omega \left\{ \nabla \cdot (\partial_t w_n \nabla w_n) - (\partial_t \nabla w_n) \cdot \nabla w_n \right\} dx dt.
	\end{equation*}
	So after using the divergence theorem together with the homogeneous boundary data, as well as noticing that $(\partial_t \nabla w_n) \cdot \nabla w_n = \frac{1}{2}\partial_t |\nabla w_n|^2$, we see that
	\begin{equation*}
	\int_0^T \zeta \int_\Omega \partial_t w_n \Delta w_n dxdt = -\iint_{Q_T}\frac{\zeta}{2}\partial_t |\nabla w_n|^2 dxdt.
	\end{equation*}
	Integrating by parts in the temporal dimension, we end up with
	\begin{equation*}
	\int_0^T \zeta \int_\Omega \partial_t w_n \Delta w_n dxdt = \iint_{Q_T} \frac{\partial_t \zeta}{2} |\nabla w_n|^2 dxdt - \frac{1}{2}\int_\Omega \left\{ |\nabla w_n(T)|^2 - |\nabla w_n(0)|^2 \right\} dx.
	\end{equation*}
	A similar inetgration by parts in the temporal dimension of the last term on the right hand side in (\ref{temporal_bound_start}) yields
	\begin{equation*}
	\iint_{Q_T} \zeta f_n \partial_t w_n dxdt = -\iint_{Q_T} \partial_t (\zeta f_n) w_n dxdt + \int_\Omega \left\{ f_n(T)w_n(T) - f_n(0)w_n(0) \right \} dx.
	\end{equation*}
	For the last term in this integral we can do the same trick as we did previously for the energy estimate. We use Cauchy-Schwarz and the inequality $\sqrt{ab} \leq \frac{c^2}{4}a + \frac{1}{c^2}b$. Then, after using a Poincaré inequality we can choose $c$ so that the integrals of $|\nabla w_n|^2$ at the temporal endpoints are cancelled. In toto we end up with
	\begin{align*}
	\iint_{Q_T} \zeta \Phi'_n(u_n) |\partial_t u_n|^2 dxdt &\leq \iint_{Q_T} \left\{ \frac{\partial_t \zeta}{2}|\nabla w_n|^2 - \partial_t(\zeta f_n) w_n \right\} dxdt \\
	&+ \frac{C}{2}\int_\Omega \left\{ f_n^2(T) - f_n^2(0) \right\} dx	\\
	&= \iint_{Q_T} \left\{ \frac{\partial_t \zeta}{2}|\nabla w_n|^2 - \partial_t(\zeta f_n) w_n \right\} dxdt \\
	&+ C \iint_{Q_T} f_n\partial_t f_n dxdt.
	\end{align*}
	Thus, by our assumption that $f$ and $\partial_t f$ are bounded, and our previous energy bound, the left hand side is bounded independent of $n$.

	In particular, we can choose $\zeta = 1$\footnote{Yikes, sorry about the kind of a fortiori argument, those are always a bit of a headache.}, and we get a bound on $\Phi'_n(u_n)|\partial_t u_n|^2$ in $L^1(Q_T)$. See now that $|\partial_t w_n	|^2 = \Phi'_n(u_n)^2|\partial_t u_n|^2$. And so, beacuse $\Phi'(u)$ is bounded for all $|u| \leq C$, we have that $\Phi'_n(u_n)$ is uniformly bounded, our latest integral bound gives us that $\partial_t w_n$ is uniformly bounded in $L^2(Q_T)$.

	\item[Using a compactness argument:] We can conclude that $\{w_n\}_{n \in \mathbb{N}}$ is a bounded sequence in $H^1(Q_T)$, so by theorem \ref{thm:Rellich} it has convergent subsequence $\{w_{n_j}\}_{j \in \mathbb{N}}$ that converges to some $w$ in $L^2(Q_T)$. The energy estimate continues to hold in the limit as
	\begin{equation}
	\iint_{Q_T}|\nabla w|^2 dxdt \leq \int_\Omega \Psi(u_0)dx + \iint_{Q_T}fw dxdt.
	\end{equation}
	Notice that since $w \in L^2(Q_T)$, we can from this inequality conclude that $w \in H^1(Q_T)$.

	\item[Convergence of $u_n$:] Recall that we defined $w_n = \Phi_n(u_n)$. Consider now
	\begin{equation*}
	\partial_t (u_{n} - u_{m}) = \Delta (w_{n} - w_{m}) + (f_{n}-f_{m}).
	\end{equation*}
	We will now do quite the same as we did when the $L^1$-contracivity for classical solutions: Let $p: \mathbb{R} \to [0,1]$ be a smooth approximation of the heaviside function, and multiply the above equation by $p(w_{n}-w_{m})$. Integrating over $\Omega$ and using integration by parts results in
	\begin{align*}
	\int_\Omega \partial_t(u_n - u_m)p(w_n-w_m)dx &= -\int_\Omega p'(w_n-w_m)|\nabla(w_n-w_m)|^2dx + \int_\Omega (f_n-f_m)p(w_n-w_m)dx \\
	&\leq \int_\Omega (f_n-f_m)_+ dx.
	\end{align*}
	This last step us justified in that for all $n$ we have $w_n \in H^1(\Omega)$ for  every $t \in [0,T]$. Letting $p$ tend towards the heaviside function and integrating over the time interval $[0,t]$ we end up with
	\begin{equation*}
	\int_\Omega (u_n - u_m)_+(t) dx \leq \int_\Omega (u_{0,n} - u_{0,m})_+dx + \int_0^t\int_\Omega (f_n - f_m)_+ dxds,
	\end{equation*}
	and as a consequence
	\begin{equation*}
	||(u_n - u_m)(t)||_{L^1(\Omega)} \leq ||u_{0,n} - u_{0,m}||_{L^1(\Omega)} + ||f_n-f_m||_{L^1(Q_t)}.
	\end{equation*}
	Integrating over time yet again and using that $Q_T$ is bounded we deduce that $\{u_n\}$ is a Cauchy sequence in $L^1(Q_T)$, and hence converges to some $u$ in $L^1$.

@@

	\item[$u$ is a weak solution:] We already know that $u \in L^1(Q_T)$, and $\Phi(u)=w \in L^1(0,T: W^{1,1}(\Omega))$\footnote{Actually, this may not be as easy to see as I've stated here. Need to use yet another estimate to show this.}. So what's left is to show that
	\begin{equation*}
	\iint_{Q_T}\left\{ \nabla\Phi(u) \cdot \nabla \eta - u\partial_t \eta - f\eta \right\}dxdt - \int_\Omega u_0\eta\big|_{t=0}dx = 0
	\end{equation*}
	for any $\eta \in C^1(\overline{Q}_T)$ that vanish on $\partial \Omega \times [0,T)$ and for $t=T$. Seeing as every $u_n$ is a weak solution to the approximate (\ref{approximate_problem}) we can subtract it from the integral above to get
	\begin{align*}
	&\iint_{Q_T}\left\{ \nabla\Phi(u) \cdot \nabla \eta - u\partial_t \eta - f\eta \right\}dxdt - \int_\Omega u_0\eta\big|_{t=0}dx \\
	= &\iint_{Q_T}\left\{ \nabla(\Phi(u)-w_n) \cdot \nabla \eta - (u-u_n)\partial_t \eta - (f-f_n)\eta \right\}dxdt - \int_\Omega (u_0-u_{0,n})\eta\big|_{t=0}dx.
	\end{align*}

	All the terms but the last are controlled by the convergences of $u_n$, $u_{0,n}$, and $f_n$.  	
\end{description}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%			SPATIAL DER. BOUND
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Control of the spatial derivative}
This bound follows nicely the $L^1$-contractivity in that its derivation hinges on a clever choice of function to multiply (\ref{GPME}) with. Before proceeding, we define  $\Psi$ as the primitive of $\Phi$ with $\Psi(0) = 0$, i.e.
\begin{equation}
\label{phi_primitive}
	\Psi(s) = \int_0^s \Phi(u) ds,
\end{equation}

@@

\begin{obs}
We can go even further and deal with the last term on the right hand side: Assuming $f,\Phi(u) \in L^2(Q_T)$, then
\begin{equation*}
\iint_{Q_T} f\Phi(u)dxdt \leq \left(\iint_{Q_T} f^2 dxdt\right)^\frac{1}{2}\left(\iint_{Q_T} \Phi(u)^2 dxdt\right)^\frac{1}{2},
\end{equation*}
by Cauchy-Schwarz'. With the inequality $\sqrt{ab} \leq \frac{1}{4c}a + cb$ for $c>0$
\begin{equation*}
\iint_{Q_T} f\Phi(u)dxdt \leq \frac{1}{4c}\iint_{Q_T}f^2dxdt + c\iint_{Q_T}\Phi(u)^2dxdt.
\end{equation*}
Lastly, because of the homogeneous Dirichlet conditions, we use the Poincaré inequality and set $c$ so that
\begin{equation}
\label{using_Poincare}
\iint_{Q_T} f\Phi(u)dxdt \leq C\iint_{Q_T}f^2dxdt + \frac{1}{2}\iint_{Q_T}|\nabla \Phi(u)|^2dxdt,
\end{equation}
and the final inequality becomes
\begin{equation}
\frac{1}{2}\iint_{Q_T}|\nabla \Phi(u)|^2dxdt + \int_\Omega \Psi(u(T))dx \leq \int_\Omega \Psi(u_0)dx + C\iint_{Q_T}f^2dxdt.
\end{equation}
\end{obs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%				TEMPORAL DER. BOUND
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Control of the temporal derivative}
A similar, but perhaps a bit more involved argument can be used to make an estimate the temporal derivative.

@@

A particular choice for $\zeta$, that's of interest is $\zeta(t) = 1$. Then we end up with the estimate
\begin{equation}
\label{temporal_bound}
\iint_{Q_T} \Phi'(u) |\partial_t u|^2 dxdt \leq \iint_{Q_T}\left\{ -\partial_t f \Phi(u)   +Cf\partial_tf \right\}dxdt, 
\end{equation}
and in addition if $\Phi'(u)$ is bounded over all relevant values of $u$, e.g. given by the $L^\infty$-bound, then what we have here is an $L^2$-estimate of $\partial_t \Phi(u)$.

@@

	\item[Control of the temporal derivative:] The derivation of (\ref{temporal_bound}) holds up in this case and we have
	\begin{equation*}
	\iint_{Q_T} \Phi'_n(u_n)|\partial_t u_n|^2 dxdt \leq \iint_{Q_T} \left\{ -\partial_t f_n \Phi_n(u_n) + Cf_n\partial_tf_n \right\} dxdt.
	\end{equation*}
	Just to reiterate, $C$ is here the constant given by Poincaré inequality, and is as such only dependent on $\Omega$.

	Now, since we have assumed $\Phi$ to be locally Lipschitz, $|\Phi_n'(u_n)| \leq K < \infty$ over the relevant values of $u_n$, we notice that $|\partial_t w_n|^2 \leq K|\Phi'_n(u_n) (\partial_t u_n)^2|$. We have assumed both $f$ and $\partial_t f$ to be bounded, so we can put a uniform bound on the right hand side, independent of $n$. We conclude that $\partial_t w_n$ is uniformly bounded in $L^2(Q_T)$.

	\item[Using a compactness argument:]From the two previous parts we have that $\{w_n\}_{n \in \mathbb{N}}$ is a bounded sequence in $H^1(Q_T)$, so by theorem \ref{thm:Rellich} it has convergent subsequence $\{w_{n_j}\}_{j \in \mathbb{N}}$ that converges to some $w$ in $L^2(Q_T)$. The energy estimate continues to hold in the limit as
	\begin{equation}
	\iint_{Q_T}|\nabla w|^2 dxdt \leq \int_\Omega \Psi(u_0)dx + \iint_{Q_T}fw dxdt.
	\end{equation}
	Notice that since $w \in L^2(Q_T)$, we can from this inequality conclude that $w \in H^1(Q_T)$.

	\item[Convergence of $u_{n_j}$:] Since all $u_{n_j}$ are classical solutions, they're also weak solutions. Therefore, when putting $u_{n_j}$ and $u_{n_k}$ into (\ref{weak}) and taking their difference, 
	\begin{align*}
	-\iint_{Q_T}(u_{n_j} - u_{n_k})\partial_t \eta dxdt &= -\iint_{Q_T}\nabla(w_{n_j} - w_{n_k})\cdot \nabla \eta dxdt \\
	&+ \int_\Omega (u_{0,n_j} - u_{0,n_k})\eta\big|_{t=0}dx \\
	&+ \iint_{Q_T}(f_{n_j} - f_{n_k})\eta dxdt 
	\end{align*}	 
	pops out. Restricting ourselves to only $\eta \in C^\infty_c(Q_T)$, and doing integration by parts on the term involving $w$, yields
	\begin{equation*}
	\bigg|\iint_{Q_T}(u_{n_j} - u_{n_k})\partial_t \eta dxdt \bigg| \leq \bigg|\iint_{Q_T} (w_{n_j} - w_{n_k})\Delta\eta dxdt \bigg| +  \bigg|\iint_{Q_T}(f_{n_j} - f_{n_k})\eta dxdt\bigg|,
	\end{equation*}
	where the term concerning the initial data has dropped out following that $\eta \big|_{t=0} = 0$. The first and the last term can both be made arbitrarily small by the convergence of $w_{n_j}$ and $f_n$ respectively. And so $\partial_t u_{n_j}$ in the weak sense is a Cauchy sequence in $L^1_{loc}(Q_T)$, and since $Q_T$ is bounded, also in $L^1(Q_T)$.

@@

	Lastly, they're weak derivatives also agree, since
	\begin{equation*}
	\iint_{Q_T}\nabla(\Phi(u) - w) \cdot \psi dxdt = -\iint_{Q_T}(\Phi(u)-w)\nabla \cdot \psi dxdt = 0 
	\end{equation*}
	for every test function $\psi$.
	 And so the energy inequality (\ref{energy_ineq}) continues to hold as
	\begin{equation}
	\iint_{Q_T}|\nabla \Phi(u)|^2dxdt + \int_\Omega \Psi(u(x,T))dx \leq \int_\Omega \Psi(u_0)dx + \int_{Q_T} f\Phi(u)dxdt.
	\end{equation}

	\item[$u$ is a weak energy solution:] We already know that $u \in L^1(Q_T)$, and $\Phi(u)=w \in L^1(0,T: W^{1,1}(\Omega))$\footnote{Actually, this may not be as easy to see as I've stated here. Need to use yet another estimate to show this.}. So what's left is to show that
	\begin{equation}
	\label{claim_weak_solution}
	\iint_{Q_T}\left\{ \nabla\Phi(u) \cdot \nabla \eta - u\partial_t \eta - f\eta \right\}dxdt - \int_\Omega u_0\eta\big|_{t=0}dx = 0
	\end{equation}
	for any $\eta \in C^1(\overline{Q}_T)$ that vanish on $\partial \Omega \times [0,T)$ and for $t=T$. Seeing as every $u_n$ is a weak solution to the approximate (\ref{approximate_problem}) we can subtract it from the integral above to get
	\begin{align*}
	&\iint_{Q_T}\left\{ \nabla\Phi(u) \cdot \nabla \eta - u\partial_t \eta - f\eta \right\}dxdt - \int_\Omega u_0\eta\big|_{t=0}dx \\
	= &\iint_{Q_T}\left\{ \nabla(\Phi(u)-w_n) \cdot \nabla \eta - (u-u_n)\partial_t \eta - (f-f_n)\eta \right\}dxdt - \int_\Omega (u_0-u_{0,n})\eta\big|_{t=0}dx.
	\end{align*}

	All the terms are controlled by the convergences of $w_n$, $u_n$, $u_{0,n}$, as long as the limit is taken along the subsequence for which we know $w_n$ converges. Thus, (\ref{claim_weak_solution}) holds, and $u$ is a weak energy solution.

	We have from before that $u\in L^1(Q_T)$ as well as $\Phi(u) \in L^2(0,T: H^1(\Omega))$, so the solution is unique.

	\item[Considering the $L^1$-contractivity:] Let now $\hat{u}$ be the weak energy solution to some auxiliary problem with intial value $\hat{u}_0$ and loading $\hat{f}$. We also let $\hat{u}_n$ be a sequence of approximations similarly constructed as those for $u$. We go by way of first showing that
	\begin{equation}
	\label{weak_solutions_l1_start}
	\int_\Omega(u(t)-\hat{u}(t))_+dx \leq \int_\Omega (u_0 + \hat{u}_0)_+dx + \int_0^t \int_\Omega (f- \hat{f})_+ dxdt. 
	\end{equation}
	With that in mind it's worth noting that for $\int_\Omega (\cdot)_+dx$ the triangle inequality is valid, so
	\begin{align*}
	\int_\Omega (u(t) - \hat{u}(t))_+dx &\leq \int_\Omega(u(t) - u_n(t))_+dx + \int_\Omega (u_n(t) - \hat{u}_n(t))_+dx \\
	&+ \int_\Omega(\hat{u}_n - \hat{u})_+dx.
	\end{align*}
	The first and last term we can make arbitrarily small, and for the middle term we use the $L^1$-contractivity for smooth solutions. Then, for every $\epsilon > 0$
	\begin{equation*}
	\int_\Omega (u(t) - \hat{u}(t))_+dx < \epsilon + \int_\Omega(u_{0,n} + \hat{u}_{0,n})_+ dx + \int_0^t \int_\Omega (f_n - \hat{f}_n)_+dxdt,
	\end{equation*}
	for sufficiently large $n$. When using the triangle inequality again
	\begin{align*}
	\int_\Omega (u(t) - \hat{u}(t))_+dx &< \epsilon + \int_\Omega(u_{0,n} - u_0)_+dx + \int_\Omega(u_0 - \hat{u}_0)_+dx \\
	&+ \int_\Omega (\hat{u}_0 - \hat{u}_{0,n})_+dx + \int_0^t \int_\Omega(f_n - f)_+dxdt\\
	&+ \int_0^t(f-\hat{f})_+dxdt + \int_0^t\int_\Omega (\hat{f}-\hat{f}_n)_+ dxdt.
	\end{align*}	
	All the term involving $n$ in the integrand can now be made arbitrarily small, so for sufficiently large $n$
	\begin{equation*}
	\int_\Omega (u(t) - \hat{u}(t))_+dx < 2\epsilon +\int_\Omega(u_0 - \hat{u}_0)_+dx + \int_0^t \int_\Omega (f-\hat{f})dxdt.
	\end{equation*}					
	Since this should hold for every $\epsilon >0 $, we must have that (\ref{weak_solutions_l1_start}) is satisfied.

	From (\ref{weak_solutions_l1_start}), we do as for (\ref{l1_contractivity}), and get
	\begin{equation*}
	||u(t) - \hat{u}(t)||_{L^1(\Omega)} \leq ||u_0 - \hat{u}_0||_{L^1(\Omega)} + ||f-\hat{f}||_{L^1(Q_t)}
	\end{equation*}
\end{description}
